---
title: "Assignment"
author: "Akash"
date: "11/07/2021"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Load pnas_expression data 

### Load pnas_expression data with rownames as gene name (ensembl_ID)

```{r}
pnasData <- read.table("pnas_expression.txt", header = T, row.names=1, stringsAsFactors=FALSE)
dim(pnasData)
head(pnasData)
```
### Remove the column 8 and keep the rows whose sum of all column not equals zero

```{r}
pnasData <- pnasData[rowSums(pnasData[,-8])>0,-8]
dim(pnasData)
head(pnasData)

```
# 2. Perform Upper quantile normalization
```{r}
pnas_uq  <- apply(pnasData, 2, function(x){quantile(pnasData[pnasData>0], 0.75)})
pnas_uq_normalized <- t(t(pnasData) / pnas_uq)
dim(pnas_uq_normalized)
head(pnas_uq_normalized)

```
# 3. Perform Deseq normalization
### 

```{r}
suppressPackageStartupMessages(library(DESeq2))
coldata <- data.frame(condition=c(rep("C",4), rep("T",3)))
dim(coldata)
head(coldata)
row.names(coldata) <- colnames(pnasData)
head(coldata)

```
### Create dataset in DSeq2
```{r}
dds <- DESeqDataSetFromMatrix(countData = pnasData, colData = coldata, design = ~ condition)
dds <- estimateSizeFactors(dds)
sizeFactors(dds)

```
### Get the normalized dataset

```{r}
pnas_deseq_normalized <- counts(dds, normalized=T)
```

# 4. Box Plot
### This is a helper function that takes data frame as input parameter and draw Box Plot using ggplot2
```{r}

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))


plot_ggboxPlot  <-  function(data, title, yName, ymax){
  p <- ggplot(data, aes(x = key, y = value)) + 
    geom_boxplot(fill = "#4271AE", colour = "#1F3552", alpha = 0.6, outlier.shape=NA)  +
    labs(x="")  +
    scale_y_continuous(name=yName, limits=c(0, ymax)) +
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5))
  
  p
}

```
### Do the Log transformation before plotting the Box plot

### Log2 transformation of raw data 
```{r }
log_pnasData <- log2(pnasData +1)
```

### convert Upper quantile normalized data (matrix) to data Frame and then do Log transformation
```{r }
pnas_uq_normalized_data  <- as.data.frame(apply(pnas_uq_normalized, 2, unlist))
log_pnas_uq_normalized_data <- log2(pnas_uq_normalized_data +1)
```

#### convert DeSeq normalized data (matrix) to data Frame and then do Log transformation
```{r }
pnas_deseq_normalized_data  <- as.data.frame(apply(pnas_deseq_normalized, 2, unlist))
log_pnas_deseq_normalized_data <- log2(pnas_deseq_normalized_data +1)

```

### Create ggplot objects 

```{r }

p_raw <- plot_ggboxPlot(gather(log_pnasData), "Before normalization", "Log2(Read counts)", 15)
p_uq <- plot_ggboxPlot(gather(log_pnas_uq_normalized_data), "After upper quartile normalization","Log2(Read counts)", 4)
p_deseq <- plot_ggboxPlot(gather(log_pnas_deseq_normalized_data), "After DESEQ normalization", "Log2(Read counts)", 15)

```


### Draw the ggplot objects in grid using cowplot
```{r }
suppressPackageStartupMessages(library(cowplot))
plot_grid(p_raw, p_uq,p_raw, p_deseq, ncol = 2, nrow = 2)
```








